---
title: "Brute Force Search"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Brute Force Search}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rnndescent)
```

If your dataset is sufficiently "small" (either in terms of number of points or
number of dimensions) you can use brute force search to find the nearest
neighbors: just calculate the distances between all pairs of points.

This has the advantages of simplicity and accuracy. But it is of course very
slow. That said as it can be easily parallelized, you can get a surprisingly
long way with it, especially if you only need the k-nearest neighbors and won't
be running a neighbor search multiple times.

Let's calculate the exact 15 nearest neighbors for the `iris` dataset:

```{r brute force knn}
iris_nbrs <- brute_force_knn(iris, k = 15)
lapply(iris_nbrs, function(x) {
  head(x[, 1:5], 3)
})
```

And here's an example of querying the odd items in the `iris` data against
the even items (i.e. the indices in `iris_nbrs` will refer to the odd items):

```{r brute force query}
iris_even <- iris[seq_len(nrow(iris)) %% 2 == 0,]
iris_odd <- iris[seq_len(nrow(iris)) %% 2 == 1,]
iris_query_nbrs <-
  brute_force_knn_query(query = iris_even,
                        reference = iris_odd,
                        k = 15)
lapply(iris_query_nbrs, function(x) {
  head(x[, 1:5], 3)
})
```
