---
title: "Querying Data"
output: 
  rmarkdown::html_vignette:
    fig_width: 4
    fig_height: 4
vignette: >
  %\VignetteIndexEntry{Querying Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bibtex
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rnndescent)
```

The usual pattern of usage for approximate nearest neighbors methods is:

1. Build some kind of index with input data.
2. Query that index with new data to find the nearest neighbors of your query.

This is how e.g. [Annoy](https://github.com/spotify/annoy) and 
[hnswlib](https://github.com/nmslib/hnswlib) work. If you want just the
k-nearest neighbors of the data you used to build the index in step 1, then you
can just pass that data as the query in step 2, but `rnndescent` provides some
specialized functions for this case that are slightly more efficient, see for
example `nnd_knn` and `rpf_knn`. Nonetheless, querying the index with the
original data can produce a more accurate result. See the [hubness](hubness.Rmd)
vignette for an example of that.

Below we will see some of the options that `rnndescent` has for querying an
index.

For convenience, I will use all the even rows of the `iris` data to build an
index, and search using the odd rows:

```{r iris split}
iris_even <- iris[seq_len(nrow(iris)) %% 2 == 0, ]
iris_odd <- iris[seq_len(nrow(iris)) %% 2 == 1, ]
```

## Brute Force

If your dataset is small enough, you can just use brute force to find the neighbors.
No index to build, no worry about how approximate the results are:

```{r brute force query}
brute_nbrs <- brute_force_knn_query(
  query = iris_odd,
  reference = iris_even,
  k = 15
)
```

The format of `brute_nbrs` is the usual k-nearest neighbors graph format, a list
of two matrices, both of dimension `(nrow(iris_odd), k)`. The first matrix, 
`idx` contains the indices of the nearest neighbors, and the second matrix,
`dist` contains the distances to those neighbors (here I'll just show the first
five results per row):


```{r brute force query results}
lapply(brute_nbrs, function(m) {
  head(m[, 1:5])
})
```

## Random Projection Forests

If you build a random projection forest with `rpf_build`, you can query it with
`rpf_knn_query`:

```{r rpf_knn_query}
rpf_index <- rpf_build(iris_even)
rpf_nbrs <- rpf_knn_query(
  query = iris_odd,
  reference = iris_even,
  forest = rpf_index,
  k = 15
)
```

See the [Random Partition Forests vignette](random-partition-forests.html) for
more.

## Graph Search

See [@dobson2023scaling] for an overview of graph search algorithms, which can
be described as a greedy beam search over a graph: to find the nearest
neighbors, you start at a candidate in the graph, find the distance from that
candidate to the query point, and update the neighbor list of your query
accordingly. If the candidate made it into the neighbor list of the query, this
seems like a promising direction to go in, so add the candidate's neighbors to
the list of candidates to explore. Repeat this until such a time as you run out
of candidates. You may want to explore the neighbors of the candidate even if it
doesn't make it onto the current neighbor list, if its distance is sufficiently
small. How much tolerance you have for this controls how much back-tracking you
do and hence how much exploration and the amount of time you spend in the
search.

`graph_knn_query` implements this search. At the very least you must provide a
`reference_graph` to search, the `reference` data that built the
`reference_graph` (so we can calculate distances), `k` the number of neighbors
you want, and of course the `query` data:

```{r graph_knn_query}
graph_nbrs <- graph_knn_query(
  query = iris_odd,
  reference = iris_even,
  reference_graph = rpf_nbrs,
  k = 15
)
```

If you aren't using the `metric = "euclidean"`, you should also provide the same
`metric` that you used to build the `reference_graph`. The default `metric` is
always `"euclidean"` for any function in `rnndescent` so it's not provided in
the examples here.

There are some other parameters you will want to tweak in any real world case
that merit some deeper discussion.

### `n_threads`

`n_threads` controls how many threads to use in the search. Be aware that
`graph_knn_query` is designed for *batch* parallelism, and each thread will be
responsible for searching a subset of the `query` points. This means that in a
streaming context, where queries to search are likely to arrive one at a time,
you won't get any speed up from using multiple threads.

### `epsilon`

`epsilon` controls how much exploration of the neighbors of a candidate to do,
as suggested by [@iwasaki2018optimization]. The default value is `0.1`, which
is also the default of the 
[NGT](https://github.com/yahoojapan/NGT/blob/main/python/README-ngtpy.md)
library. The larger the value, the more back-tracking is permitted. The exact
meaning of the value is related to how large a distance is considered "close
enough" the current neighbor list of the query to be worth exploring.

`epsilon = 0.1` means that the query-candidate distance is allowed to be 10%
larger than the largest distance in the neighbor list. If you set `epsilon =
0.2`, for example, then the query-candidate distance is allowed to be 20% higher
than the largest distance in the neighbor list and so on. If you set 
`epsilon = 0` then you get a pure greedy search.

It's hard to give a general rule for what value to set, because it's highly
dependent on the distribution of distances in the dataset and that is
determined by the distance metric and the dimensionality of the data itself.
I recommend leaving this as the default, and only modifying it if you find that
the search is unreasonably slow (in which case make `epsilon` smaller) or
unreasonably inaccurate (in which case make `epsilon` larger). Yes, not very
helpful I know. In the benchmarking done in [@dobson2023scaling] using a similar
back-tracking method, `epsilon = 0.25` was the maximum value used and in
[@wang2021comprehensive] `epsilon = 0.1` was used.
 
### `init`

This controls how the search is initialized. If you don't provide this, then
`k` random neighbors per item in `query` will be generated for you.
 
#### Neighbor Graph Input
 
You may provide your own input for this. It should be in the neighbor graph
format, i.e. a list of two matrices, `idx` and `dist`, as described above. Make
sure that the `dist` matrix contains the distances using the same `metric` you
will use in the search.

#### Neighbor Indices Only

In fact, the `dist` matrix is optional. If you only provide the `idx` matrix,
then the `dist` matrix will be calculated for you. If the `dist` matrix is
already available to you and it was generated by `rnndescent` then there is no
reason *not* to use it, but you could have neighbors that come from:

* another nearest neighbor package and for some reason you don't have the distance
* or you have indices from a different metric that you nonetheless believe are
a good guess for the "real" metric.

A case where this might be worth experimenting with could be if you can cheaply
binarize your input data, i.e. convert it to 0/1 then to `FALSE`/`TRUE`: you 
could then use the `hamming` metric or another binary-specialized metric on that
input data. Even a brute force search can be very fast on this data. This could be
a good way to get a good guess for the real data.

This is a very contrived example with `iris`, but let's do it anyway:

```{r iris binarize}
numeric_iris <- iris[, sapply(iris, is.numeric)]
logical_iris <- sweep(numeric_iris, 2, colMeans(numeric_iris), ">")
logical_iris_even <- logical_iris[seq_len(nrow(logical_iris)) %% 2 == 0, ]
logical_iris_odd <- logical_iris[seq_len(nrow(logical_iris)) %% 2 == 1, ]
head(logical_iris_even)
```

Do a brute force search on the binarized data:

```{r iris binarize brute force}
iris_logical_brute_nbrs <- brute_force_knn_query(
  query = logical_iris_odd,
  reference = logical_iris_even,
  k = 15,
  metric = "hamming"
)
```

Then pass the indices of the brute force search to `graph_knn_query`, which
will generate the Euclidean distances for you:

```{r iris search with binary init}
graph_nbrs <- graph_knn_query(
  query = iris_odd,
  reference = iris_even,
  reference_graph = rpf_nbrs,
  init = iris_logical_brute_nbrs$idx,
  k = 15
)
```

Whether this is worth doing all depends on whether the time taken to binarize
the data followed by the initial search on the binary data (it doesn't have to
be brute force) gives you a good enough guess to save time in the "real" search
with `graph_knn_query`.

#### Forest initialization

If you have previously built an RP Forest with the data you may also use that
to initialize the query. We can re-use `rpf_index` here.

```{r iris forest init}
forest_init_nbrs <- graph_knn_query(
  query = iris_odd,
  reference = iris_even,
  reference_graph = rpf_nbrs,
  init = rpf_index,
  k = 15
)
```

In general, the RP forest initialization is likely to be a better initial guess
than random, but in terms of a speed/accuracy trade-off, using a large forest
may not be the best choice. You may want to use `rpf_filter` to reduce the size
of the forest before using it as an initial guess. In the
[PyNNDescent](https://github.com/lmcinnes/pynndescent) Python package that
`rnndescent` is based on, only one tree is used for initializing query results.

## Preparing the Search Graph

In all the examples so far, we have used the k-nearest neighbors graph as the
`reference_graph` input to `graph_knn_query`. Is this actually a good idea?
Probably not! There is no guarantee that all the items in the original dataset
can actually be reached via the k-nearest neighbors graph. Some nodes just
aren't very popular and may not be in the neighbor list of *any* other item.
That means you can never reach them via the k-nearest neighbors graph, no matter
how thoroughly you search it. 

We can solve this problem by reversing all the edges in the graph and adding
them to the graph. So if you can get to item `i` from item `j`, you can now get
to item `j` from item `i`. This solves one problem but adds some more which is
that just like some items are very unpopular, other items might be very popular
and appear often in the neighbor list of other items. Having a large number of
these edges in the graph can make the search very slow. We therefore need
to prune some of these edges.

`prepare_search_graph` is a function that will take a k-nearest neighbor graph
and add edges to it to make it more useful for a search. The procedure is based on
the process described in [@harwood2016fanng] and consists of:

1. Reversing all the edges in the graph.
1. "Diversifying" the graph by "occlusion pruning". This considers triplets of
points, and removes long edges which are probably redundant. For an item $i$
with neighbors $p$ and $q$ if the distances $d_{pq} \lt d_{ip}$ i.e. the
neighbors are closer to each other than they are to $i$, then it is said that
$p$ occludes $q$ and we don't need both edges $i \rightarrow p$ and
$i \rightarrow q$ -- it's likely that $q$ is in the neighbor list of $p$ or vice
versa, so it's unlikely that we are doing any harm by getting rid of 
$i \rightarrow p$.
1. After occlusion pruning, if any item still has an excessive number of edges,
the longest edges are removed until the number of edges is below the threshold.

To control all this pruning the following parameters are available:

* `diversify_prob`: the probability of a neighbor being removed if it is found
  to be an "occlusion". This should take a value between `0` (no diversification)
  and `1` (remove as many edges as possible). The default is `1.0`.
* `pruning_degree_multiplier`: controls how many edges to remove after the
  occlusion pruning relative to the number of neighbors in the original nearest
  neighbor graph. The default is `1.5` which means to allow as many as 50%
  more edge than the original graph. So if the input graph was for `k = 15`,
  each item in the search graph will have at most `15 * 1.5 = 22` edges.
  
Let's see how this works on the `iris` neighbors:

```{r iris search graph}
set.seed(42)
iris_search_graph <- prepare_search_graph(
  data = iris_even,
  graph = rpf_nbrs,
  diversify_prob = 0.1,
  pruning_degree_multiplier = 1.5
)
```

Because the returned search graph can contain different number of edges per item,
the neighbor graph format isn't suitable. Instead you get back a sparse matrix,
specifically a `dgCMatrix`. Here's a histogram of how the edges are distributed:

```{r iris search graph histogram}
search_graph_edges <- diff(iris_search_graph@p)
hist(search_graph_edges,
  main = "Distribution of search graph edges", xlab = "# edges"
)
range(search_graph_edges)
```

So most items have around about `k = 15` edges just like the nearest neighbor
graph. But some have have the maximum number of edges and few have only 10
edges.


```{r iris use a search graph}
search_nbrs <- graph_knn_query(
  query = iris_odd,
  reference = iris_even,
  reference_graph = iris_search_graph,
  init = rpf_index,
  k = 15
)
```

## References
