---
title: "rnndescent"
output: 
  rmarkdown::html_vignette:
    fig_width: 4
    fig_height: 4
vignette: >
  %\VignetteIndexEntry{rnndescent}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bibtex
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rnndescent)
```

Nearest Neighbor Descent (NND) [@dong2011efficient] is affected by hubness 
[@bratic2019influence]: this is when some items in a dataset appear as a near 
neighbor of other points very frequently. This can result in reduced accuracy
of the approximate nearest neighbor graph produced by NND and may be an
intrinsic problem in high dimensional datasets (although see [@low2013hubness]
for a dissenting view).

In this vignette we will use synthetic data to explore the issue, and use the
k-occurrences of a neighbor graph to identify when NND is at risk of producing
less accurate results. We will also look at some ways to ameliorate the effects
of hubness.

First, to control the pseudo-random number generation:

```{r set seed}
set.seed(42)
```

Now let's create some Gaussian data to test with. First, a low-dimensional
example:

```{r 2D Gaussian}
n_points <- 1000
low_dim <- 2
g2d <- matrix(rnorm(n_points * low_dim), ncol = low_dim)
```

In this vignette, we are interested in the 15 nearest neighbors. To get the 
exact nearest neighbors, we use the `brute_force_knn` function with `k = 15`:

```{r brute force 2D results}
g2d_nnbf <- brute_force_knn(g2d, k = 15, metric = "euclidean")
```

This will act as our "ground truth" and we will compare how well NND does. To
use NND to find the approximate nearest neighbors, we use the `nnd_knn` 
function:

```{r nearest neighbor descent 2D results}
g2d_nnd <- nnd_knn(g2d, k = 15, metric = "euclidean")
```

To calculate the accuracy of NND we will produce a vector where each element
contains the accuracy with which the exact 15-nearest neighbors were reproduced
for each item in the data: `0` means that none of the exact 15-nearest neighbors
were in the list of 15-nearest neighbors that NND found, and `1` means they all
were.

To get the vector of accuracies, here is the `nn_accuracyv` function:

```{r nn accuracy vector}
nn_accuracyv <-
  function(idx,
           ref_idx,
           k = NULL,
           include_self = TRUE,
           verbose = TRUE) {
    if (is.list(idx)) {
      idx <- idx$idx
    }
    if (is.list(ref_idx)) {
      ref_idx <- ref_idx$idx
    }

    if (is.null(k)) {
      k <- min(ncol(idx), ncol(ref_idx))
    }

    if (ncol(ref_idx) < k) {
      stop("Not enough columns in ref_idx for k = ", k)
    }

    n <- nrow(idx)
    if (nrow(ref_idx) != n) {
      stop("Not enough rows in ref_idx")
    }

    nbr_start <- 1
    nbr_end <- k

    ref_start <- nbr_start
    ref_end <- nbr_end
    if (!include_self) {
      ref_start <- ref_start + 1
      if (nbr_end < ncol(ref_idx)) {
        ref_end <- ref_end + 1
      } else {
        nbr_end <- nbr_end - 1
      }
    }

    nbr_range <- nbr_start:nbr_end
    ref_range <- ref_start:ref_end

    total_intersect <- rep(0, times = n)
    for (i in 1:n) {
      total_intersect[i] <-
        length(intersect(idx[i, nbr_range], ref_idx[i, ref_range]))
    }

    total_intersect / k
  }
```

`nn_accuracy` is the average accuracy over the whole dataset:

```{r nn dataset accuracy}
nn_accuracy <-
  function(idx,
           ref_idx,
           k = NULL,
           include_self = TRUE,
           verbose = TRUE) {
    vec <- nn_accuracyv(idx, ref_idx, k, include_self, verbose)
    sum(vec) / length(vec)
  }
```

With low-dimensional data, nearest neighbor descent does very well:

```{r 2D nnd accuracy}
nn_accuracy(g2d_nnbf, g2d_nnd, k = 15)
```

Now let's see what happens with a high-dimensional (1000 features):

```{r 1000D Gaussian}
hi_dim <- 1000
g1000d <- matrix(rnorm(n_points * hi_dim), ncol = hi_dim)
```

Again we will use brute force to generate the true nearest neighbors.

```{r brute force 1000D results}
g1000d_nnbf <- brute_force_knn(g1000d, k = 15, metric = "euclidean")
```

Let's do NND on the high dimensional data...

```{r nearest neighbor descent 1000D results}
g1000d_nnd <- nnd_knn(g1000d, k = 15, metric = "euclidean")
```

...and see how well it does:

```{r 1000D nnd accuracy}
nn_accuracy(g1000d_nnbf, g1000d_nnd, k = 15)
```

Still OK, but not as good as we might like.

## Comparing Low- and High-Dimensional Nearest Neighbors

Let's look at the distribution of nearest neighbor distances in high and low
dimensions (for easier, comparison, I have normalized them with respect to the
largest distance)

```{r NN distances distribution}
hist(g2d_nnbf$dist[, -1] / max(g2d_nnbf$dist[, -1]), xlab = "distances", main = "2D 15-NN")
hist(g1000d_nnbf$dist[, -1] / max(g1000d_nnbf$dist[, -1]), xlab = "distances", main = "1000D 15-NN")
```

Compared to low dimensional data, we can see that the high dimensional distances
are distributed around a higher distance as well as more symmetric in their 
distribution.

Here are the distribution of the neighbor distances in the high-dimensional case
for the neighbors found by NND:

```{r NND distances distribution}
hist(g1000d_nnd$dist[, -1] / max(g1000d_nnd$dist[, -1]),
  xlab = "distances",
  main = "1000D 15-NND"
)
```

Pretty much indistinguishable from the exact results, so it seems like there
isn't an obvious diagnostic from the distances themselves.

Is the distribution of the errors in the NND results uniform or are some items
neighborhoods noticeably better predicted than others? This function will
calculate a vector of the relative RMS error between two sets of neighbors in
terms of distances:

```{r distance relative RMS error}
nn_rrmsev <- function(nn, ref) {
  n <- ncol(ref$dist) - 1
  sqrt(apply((nn$dist[, -1] - ref$dist[, -1])^2 / n, 1, sum) /
    apply(ref$dist[, -1]^2, 1, sum))
}
```

We don't include the first nearest neighbor distances, as these are invariably
the self distances which leads to an uninteresting number of zero error results.
This measure of error is a bit less strict than `nn_accuracyv` as a neighbor
that is outside the true kNN, but which has a comparable distance, will be
penalized a less harshly than a more distant point.

Here's a histogram of RRMS distance errors:

```{r histogram of distance difference}
g1000d_rrmse <- nn_rrmsev(g1000d_nnd, g1000d_nnbf)
hist(g1000d_rrmse,
  main = "1000D distance error",
  xlab = "Relative RMS error"
)
```

None of the relative errors are actually that large, so if you only care about
the value of kth nearest neighbor distances, then even in the 1000D case, we still
get decent results in this case. We can also see that there is a clear distribution
of errors, where an appreciable number of items have zero RRMS distance errors,
but there a few items which have the largest error.

Here is a histograms of the accuracies:

```{r hist 1000D}
g1000d_nnd_acc <- nn_accuracyv(g1000d_nnbf, g1000d_nnd, k = 15)
hist(g1000d_nnd_acc,
  main = "1000D accuracy",
  xlab = "accuracy",
  xlim = c(0, 1)
)
```

This shows a similar pattern: some items have very high accuracy and some have
noticeably worse accuracy than average. For completeness, here is the relationship between 
accuracy and RRMSE:


```{r rrmse vs acc}
plot(
  g1000d_nnd_acc,
  g1000d_rrmse,
  main = "RRMSE vs accuracy",
  xlab = "accuracy",
  ylab = "RRMSE"
)
```

Nothing very surprising here: there's a fairly consistent spread of RRMSE for
a given accuracy.

So whether you use an error in the distance or accuracy to measure how well the
approximate nearest neighbors method is working, at least in this case, a high
dimensional dataset affects some items more than others.

## Detecting Hubness

[@radovanovic2010hubs] discusses a technique for detecting hubness: look for 
items that appear very frequently in the k-nearest neighbor graph. The `k_occur`
function counts "k-occurrences" of each item in a dataset, i.e. a count of the
number of times an item appears in the k-nearest neighbor graph. You can also
see it as reversing the direction of the edges in the k-nearest neighbor graph
and then counting the in-degree of each item.

If the distribution of neighbors was entirely uniform we would expect to see
each item appear $k$ times. If there are hubs then the k-occurrence could get
as large as the size of the dataset, $N$. An item which appears in the neighbor
graph fewer than $k$ times could be termed an "antihub". Our definition of a 
neighbor of an item always includes the item itself, so we would expect the 
minimum $k$-occurrence to be $1$. 

Also, because there are always only $Nk$ edges in a $k$-nearest neighbor graph,
if an item appears more than the expected amount this implies that other items
must be under-represented. Practically speaking, there are always going to be
items with a larger $k$-occurrence than expected and hence some with a lower
$k$-occurrence, so hubness or anti-hubness is more a case of deciding on a
cut-off after which the presence of an item with a lot of neighbors starts
causing you problems, which is going to be dependent on what you are planning to
do with the neighbor graph (and probably the number of neighbors you want).

### k-occurrence in the 2D case

First, let's look at the 2D case using the exact k-nearest neighbors:

```{r 2D k-occurrences}
g2d_bfko <- k_occur(g2d_nnbf, k = 15)
summary(g2d_bfko)
```

The mean average of the $k$-occurrence is never helpful: as noted above there
are always $Nk$ edges in the neighbor graph, so the mean $k$-occurrence is always
$k$. However the other descriptions of the distribution are informative. The
median $k$-occurrence is also `15`, which is a good sign, and the values at 25%
and 75% aren't too different other. The maximum $k$-occurrence is less than
$2k$. The minimum value is `1` which means there are anti-hubs in the dataset,
but:

```{r number of anti-hubs in 2D case}
sum(g2d_bfko == 1)
```

there is only one anti-hub in this dataset. Here's a histogram of the 
k-occurrences:

```{r 2D k-occurrence histogram}
hist(g2d_bfko, main = "2D 15-NN", xlab = "k-occurrences")
```

This unremarkable-looking distribution is a visual indication of a dataset 
without a lot of hubness and anti-hubs lurking to cause problems for nearest
neighbor descent.

### k-occurrence in the 1000D case

Here's what the k-occurrence histogram looks like for the high dimensional case:

```{r 1000D k-occurrences}
g1000d_bfko <- k_occur(g1000d_nnbf$idx, k = 15)
hist(g1000d_bfko, main = "1000D 15-NN", xlab = "k-occurrences")
```

The differences are pretty stark. The first thing to notice is the x-axis. In
the 2D case, the maximum k-occurrence was ~20. For the 1000D we are looking at
~300. It's hard to see any details, so let's zoom in on the same region as the
2D case by clipping any k-occurrence larger than the largest 2D k-occurrence:

```{r zoomed k-occurrences}
hist(pmin(g1000d_bfko, max(g2d_bfko)),
  main = "1000D 15-NN zoomed",
  xlab = "k-occurrences"
)
```

It's a very different distribution to the 2D case: we have a large number of
anti-hubs and a noticeable number of hubs. There's certainly no peak at a 
k-occurrence of 15. Comparing the numerical summary with the 2D case is 
instructive:

```{r numerical summary of 1000D k-occurrence distribution}
summary(g1000d_bfko)
```

Again, here's a good reminder that the mean k-occurrence is of no value. The
median k-occurrence immediately communicates the difference between the 2D
case. We can also see that the maximum k-occurrence means that there is one 
point which is considered a close neighbor of over one third of the dataset.

How many anti-hubs are there?

```{r number of anti-hubs in the 1000D case}
sum(g1000d_bfko == 1)
```

A quarter of the dataset does not appear as a neighbor of any other point. This
has serious implications for using a neighbor graph for certain purposes: you
cannot reach a quarter of the dataset by starting at an arbitrary point in the
graph and following neighbors.

This also might point to why nearest neighbor descent has trouble with this high
dimensional case: if we rely on points turning up as a neighbors of other points
in order to introduce them to potential neighbors, the fact that so many of the
points in this dataset aren't anyone's actual neighbors would suggest they are
unlikely to get involved in the local join procedure as much as other points.

### k-occurrence as a diagnostic of NND failure

We have now shown that we can use k-occurrences on the exact nearest neighbors
of low and high dimensional data to detect the existence of hubs, which in turn
might lead us to suspect that the approximate nearest neighbors found by nearest
neighbor descent may not be very accurate. But that's not a very useful
diagnostic because if we have the exact neighbors we don't need to run NND in
the first place. But even if the approximate nearest neighbor graph produced by
NND isn't highly accurate, does it still show similar characteristics of
hubness?

```{r 1000D NND k-occurences}
g1000d_nndko <- k_occur(g1000d_nnd$idx, k = 15)
hist(g1000d_nndko, main = "1000D 15-NND", xlab = "k-occurrences")
```

That seems similar to the true results, and zooming in like we did with the
exact results:

```{r zoomed NND k-occurrences}
hist(pmin(g1000d_nndko, max(g2d_bfko)),
  main = "1000D 15-NND zoomed",
  xlab = "k-occurrences"
)
```

Visually this looks a lot like the distribution of the exact results. Next, the 
numerical summary:


```{r 1000D NND k-occurences numeric summary}
summary(g1000d_nndko)
sum(g1000d_nndko == 1)
```

Quantitatively, this also tracks the exact results: the median k-occurrence
is much smaller than $k$, there is a hub with a very large number of neighbors
(larger than in the exact case but to a similar degree) and a similar number
of anti-hubs.

So this suggests a way to diagnose if the nearest neighbor descent routine may
have low accuracy: look at the distribution of the k-occurrences of the
resulting approximate nearest neighbor graph (or even just the maximum value). A
value that is $\gg k$ may mean a reduced accuracy. Of course, this isn't
foolproof, because even if NND did a perfect job then we would still get these
sorts of values, but it's a starting point.

Taking the distribution of k-occurrences as a whole, the approximate results
seem to track the exact results fairly well, but as we have seen, the errors
in the approximate results are not uniformly distributed across the data. So
let's see how well the NND k-occurrences "predict" the exact results:

```{r approximate vs true 1000D k-occurrence}
plot(g1000d_nndko, g1000d_bfko,
  xlab = "approximate", ylab = "exact",
  xlim = c(0, max(g1000d_nndko, g1000d_bfko)),
  ylim = c(0, max(g1000d_nndko, g1000d_bfko)),
  main = "1000D k-occ"
)
abline(a = 0, b = 1)
cor(g1000d_nndko, g1000d_bfko, method = "pearson")
```

The overall relationship seems strong. The line on the plot is x=y, so we can
see that at high values of the k-occurrence the NND results tend to 
over-estimate the k-occurrence, but these are such large values that this hardly
matters, and there is no ambiguity over which nodes are most hub-like.

Zooming in to lower values of the k-occurrence:

```{r zoomed approximate vs true 1000D k-occurrence}
plot(g1000d_nndko, g1000d_bfko,
  xlab = "approximate", ylab = "exact",
  xlim = c(0, max(g2d_bfko)),
  ylim = c(0, max(g2d_bfko)),
  main = "1000D low k-occ"
)
abline(a = 0, b = 1)
```

here it seems that there is a tendency to over-estimate the k-occurrence. 
Anti-hubs are also not perfectly identified, but there are no true anti-hubs
which appear more than a small number of times in the approximate neighbor
graph.

### Detecting Poorly Predicted Neighbors

We've seen that some objects have their neighbors predicted better than others.
Based on everything we've seen so far about k-occurrences and NND, it would be
reasonable to wonder: are the items in a dataset with poorly predicted neighbors
the anti-hubs (predicted or exact)? This would at least give us some way of 
detecting those items that were likely to have low accuracy neighborhoods:
perhaps they could be treated specially (or by some other algorithm).

Here's a plot of the accuracy against the k-occurrences of the NND neighbors:

```{r predicting accuracy with NND k-occurrence}
plot(g1000d_nndko, g1000d_nnd_acc,
  xlab = "NND k-occ", ylab = "accuracy",
  xlim = c(0, max(g1000d_nndko, g1000d_bfko)),
  main = "1000D acc vs NND k-occ"
)
```

So the answer to the question is "not really", but there *is* a trend. The empty
space in the lower right of the plot indicates that items with a large
k-occurrence (hubs) are very well predicted. And above a k-occurrence of 150, we
are guaranteed to perfectly predict the neighborhood of an item. However, at the
other end of the k-occurrence spectrum, we can see that while the lower bound on
the predicted accuracy does plummet as the k-occurrence is reduced, some
anti-hubs actually do have their neighborhoods very accurately predicted too.

Unfortunately this means that k-occurrence is a bit too rough to use to predict
poorly-predicted items. Let's say that we wanted to get all the items where the
neighborhood was less than 90% accurate:

```{r proportion of items with < 90% accuracy}
sum(g1000d_nnd_acc < 0.9)
```

That's already quite a lot of items: about three-quarters of the entire dataset.
What is the largest k-occurrence for an item in the dataset with that accuracy
threshold? 

```{r max k-occurrence for lower accuracy items}
max(g1000d_nndko[g1000d_nnd_acc < 0.9])
```

Then, to guarantee that we had found all the items that might be poorly
predicted, we would need to filter out every item that had a k-occurrence
smaller than that value, even though we know that some of them are
well-predicted:

```{r how many items}
sum(g1000d_nndko <= max(g1000d_nndko[g1000d_nnd_acc < 0.9]))
```

That's most of the dataset. If we dropped the threshold to 80 accuracy, does it
help?

```{r how many items at a lower accuracy thresold}
sum(g1000d_nndko <= max(g1000d_nndko[g1000d_nnd_acc < 0.8]))
```

A bit, but it's still a substantial majority of the dataset. So whatever we
decided to do with these items we wouldn't be saving a huge amount of effort.

So much for that idea. What this suggests is not that we *can't* improve results
here, just that the effort of identifying individual points to filter out,
treat differently and then merge back into the final neighbor graph means that
just reprocessing the entire dataset in a different way is likely to be a
competitive solution.

### Detecting Problems Early

Back to looking at the k-occurrence distribution as a whole: we can see that the
converged NND results, despite not being 100% accurate do a good job at
expressing the hubness of the underlying data. How converged do the results
need to be? What if we think of NND as a tool for identifying hubness in
datasets as a whole rather than for accurate approximate nearest neighbor
graphs? Could a much less unconverged NND graph, while obviously being even
less accurate, still correctly identify a dataset as having hubs?

To test this, let's run the NND method for only one iteration and get the
k-occurrences that result:

```{r unconverged NND}
g1000d_nnd_iter1 <- nnd_knn(g1000d, k = 15, metric = "euclidean", n_iters = 1)
g1000d_nndkoi1 <- k_occur(g1000d_nnd_iter1$idx, k = 15)
```

How accurate are these results?

```{r unconverged NND accuracy}
nn_accuracy(g1000d_nnbf, g1000d_nnd_iter1, k = 15)
```

Ok, I think we can all agree we do *not* have an accurate neighbor graph. But
let's take a look at the k-occurrence distribution:

```{r unconverged ko distribution}
hist(g1000d_nndkoi1, main = "1000D 15-NND (1 iter)", xlab = "k-occurrences")
```

Looking familiar. Zooming in...

```{r unconverged ko distribution zoomed}
hist(pmin(g1000d_nndkoi1, max(g2d_bfko)),
  main = "1000D 15-NND (1 iter, zoomed)",
  xlab = "k-occurrences"
)
```

The distribution is at least similar to the converged version. Taking a look
at some numbers:

```{r unconverged ko distribution numerical summary}
summary(g1000d_nndkoi1)
sum(g1000d_nndkoi1 == 1)
```

Compared to the converged (or exact) distribution, the median k-occurrence is
not as low, the object with the largest k-occurrence, while large ($ > 10k$,
which seems like a good threshold to be concerned about the presence of hubs) is
not as large, and there are fewer objects which are anti-hubs.

At least for this dataset, hubness can be qualitatively detected with even
a very inaccurate neighbor graph. What about datasets that don't contain hubs?
Let's just check that what we are seeing is not an artifact of unconverged
nearest neighbor descent, by running through the same procedure with the 2D
dataset:

```{r unconverged 2D NND}
g2d_nnd_iter1 <- nnd_knn(g2d, k = 15, metric = "euclidean", n_iters = 1)
g2d_nndkoi1 <- k_occur(g2d_nnd_iter1$idx, k = 15)
hist(g2d_nndkoi1, main = "2D 15-NND (1 iter)", xlab = "k-occurrences")
summary(g2d_nndkoi1)
sum(g2d_nndkoi1 == 1)
nn_accuracy(g2d_nnbf, g2d_nnd_iter1, k = 15)
```

We can see that the neighbor graph is also not very accurate after 1 iteration
in the 2D case, but the distribution of k-occurrences also qualitatively 
resembles the exact result. This time, compared to the exact results there are
slightly more anti-hubs and the maximum k-occurrence is increased, so the trends
are slightly reversed compared to the 1000D data.

For at least qualitative identification of hubness, then, one iteration of
nearest neighbor descent might be enough.

## Improving accuracy

We know that nearest neighbor descent (at least with typical settings) may not
give highly accurate results in high dimensions. And with the help of
k-occurrences, we can even detect that it might be happening. But what can we do
about it?

### Use More Neighbors

One simple (slightly expensive) way is to keep more neighbors in the
calculation. For example, double the number of neighbors to `30`, then get the
top-15 accuracy:

```{r NND 1000D truncated 30}
g1000d_nnd_k30 <- nnd_knn(g1000d, k = 30, metric = "euclidean")
nn_accuracy(g1000d_nnbf, g1000d_nnd_k30, k = 15)
```

That's a big improvement, but increasing `k` in this way can be quite expensive
in terms of run time.

### Merging Multiple Independent Results

What about taking advantage of the stochastic nature of the algorithm? If the
results are sufficiently diverse between runs of NND, then we could generate two
graphs from two separate runs, and then merge the the results.

Let's repeat NND and see what the accuracy of this new result is like.

```{r NND 1000D repeat}
g1000d_nnd_rep <- nnd_knn(g1000d, k = 15, metric = "euclidean")
nn_accuracy(g1000d_nnbf, g1000d_nnd_rep, k = 15)
```

That's similar to the first run. That's re-assuring in the sense that the
variance of the accuracy doesn't seem to be that high between one run to the
next. But hopefully that doesn't also mean that NND is producing a very similar
neighbor graph each time, in which case merging them won't be very helpful. Time
to find out:

```{r merge}
g1000d_nnd_merge <- merge_knn(g1000d_nnd, g1000d_nnd_rep)
nn_accuracy(g1000d_nnbf, g1000d_nnd_merge, k = 15)
```

That's a big improvement. So it does seem like there is some diversity in the
results. 

```{r NND 100D compare accuracy}
g1000d_nnd_rep_acc <- nn_accuracyv(g1000d_nnbf, g1000d_nnd_rep, k = 15)
plot(
  g1000d_nnd_acc,
  g1000d_nnd_rep_acc,
  main = "1000D NND accuracy comparison",
  xlab = "accuracy run 1",
  ylab = "accuracy run 2"
)
cor(g1000d_nnd_acc, g1000d_nnd_rep_acc)
```

Despite the similar overall accuracies, there's quite a large variance between
runs in terms of which items have accurate neighborhoods.

So there might be some scope for improving the results by merging different runs,
especially if you can run the individual NND routines in parallel.

### Using a Search Graph

Practically, the simplest way to improve results with `rnndescent` is to convert
the neighbor graph into a search graph, and then query it with the original 
data.

First, the preparation step:

```{r prepare graph}
g1000d_search_graph <-
  prepare_search_graph(
    data = g1000d,
    graph = g1000d_nnd,
    metric = "euclidean",
    diversify_prob = 1,
    pruning_degree_multiplier = 1.5
  )
```

This augments the neighbor graph with the reversed edges of the neighbor graph,
so that if $i$ is one of the nearest neighbors of $j$, we guarantee that $j$
is also considered a near neighbor $i$. This ameliorates the issue of anti-hubs
because all $k$ neighbors of an anti-hub now have it in their neighbor list.

The downside of including all reversed edges in the neighbor graph is that the
neighbor list of a hub is now going to be very large as it consists of the $k$
nearest neighbors of the hub and then all the items that consider the hub a
near neighbor, which by definition is a lot. This can make the search graph
inefficient, as a disproportionate amount of time will be spent searching
neighbors of the hub. The `diversify_prob` and `pruning_degree_multiplier` 
parameters are used to reduce back down the out-degree of each node (the number
of out-going edges). This results in objects with a varying number
of neighbors, in this case to a maximum of 22. This is about 50% larger than
`k = 15` to account for the introduction of the reverse edges. Anti-hubs can be
reintroduced due to the edge reduction, but hopefully the distribution of
edges is a bit more equitable.

Here is a summary and histogram of the k-occurrences of the search graph:

```{r histogram of k-occurrences of search graph}
g1000d_sgko <- k_occur(g1000d_search_graph)
hist(g1000d_sgko, main = "search graph k-occurrences", xlab = "k-occurrences")
summary(g1000d_sgko)
sum(g1000d_sgko == 1)
```

This is not *quite* as skewed as the neighbor graph, but there is still a lot of
room for improvement.

At any rate, with the search graph in hand, we can now search it using our
original data as a query:

```{r search with prepared graph}
g1000d_search <-
  graph_knn_query(
    query = g1000d,
    reference = g1000d,
    reference_graph = g1000d_search_graph,
    k = 15,
    metric = "euclidean",
    init = g1000d_nnd,
    epsilon = 0.1
  )
```

Are the results improved?

```{r accuracy with search graph}
nn_accuracy(g1000d_nnbf, g1000d_search, k = 15)
```

Yes, the accuracy is now nearly perfect. The disadvantages of the search graph
approach for building a neighbor graph is that it is less efficient than NND:
`graph_knn_query` must assume that the `query` data is entirely different to the
`reference` data. The advantage is that we can make use of reverse edges and,
more importantly, back-tracking (controlled via the `epsilon` parameter), which 
seems to make the difference in this example.

The procedure above is the recommended practice of using `graph_knn_query` with
a search graph generated from the neighbor graph. You are not required to use
a search graph as the argument to the `reference_graph` parameter. Here is the
back-tracking search using the neighbor graph directly and everything else the
same:

```{r search with neighbor graph}
g1000d_nnd_search <-
  graph_knn_query(
    query = g1000d,
    reference = g1000d,
    reference_graph = g1000d_nnd,
    k = 15,
    metric = "euclidean",
    init = g1000d_nnd,
    epsilon = 0.1
  )
nn_accuracy(g1000d_nnbf, g1000d_nnd_search, k = 15)
```

Accuracies are nearly as good. You can save even more time by turning off
back-tracking (`epsilon = 0`):

```{r search with neighbor graph, no back-tracking}
g1000d_nnd_search0 <-
  graph_knn_query(
    query = g1000d,
    reference = g1000d,
    reference_graph = g1000d_nnd,
    k = 15,
    metric = "euclidean",
    init = g1000d_nnd,
    epsilon = 0
  )
nn_accuracy(g1000d_nnbf, g1000d_nnd_search0, k = 15)
```

but accuracies are now noticeably less improved. Using the search graph without
back-tracking gives slightly better accuracies:

```{r search with search graph, no back-tracking}
g1000d_search0 <-
  graph_knn_query(
    query = g1000d,
    reference = g1000d,
    reference_graph = g1000d_search_graph,
    k = 15,
    metric = "euclidean",
    init = g1000d_nnd,
    epsilon = 0
  )
nn_accuracy(g1000d_nnbf, g1000d_search0, k = 15)
```

but it seems like some sort of back-tracking is to be recommended with this
approach.

### Local Scaling

In the context of spectral clustering, [@NIPS2004_40173ea4] suggested scaling
nearest neighbor distances to reflect the local statistics of the neighborhoods.
This would allow the idea of what a large and a short distance was to vary
depending on the local density around each point. The scaling suggested was:

$$\hat{d}_{ij}^2 = \frac{d_{ij}^2}{\sigma_i\sigma_j}$$

where $\hat{d}_{ij}$ is the scaled distance between points $i$ and $j$, 
$d_{ij}$ is the original distance, and $\sigma_i$ is the local scale associated
with point $i$. To define $\sigma_i$, Zelnik-Manor and Perona suggest using the
distance to a nearest neighbor. In their paper they got good results by using
the distance to the seventh-nearest neighbor.

[@schnitzer2012local] suggested using $\hat{d}_{ij}$ instead of $d_{ij}$ in
nearest neighbor calculations as a way to reduce the effect of hubs. Because
hubs have more neighbors, the distance to a specific neighbor will be smaller
relative to non-hubs. Hence $\sigma_i$ will be smaller for hubs, which will lead
to larger $\hat{d}_{ij}$ compared to non-hubs. Therefore using $\hat{d}_{ij}$
acts to "penalize" hubs when competing with non-hubs to appear in a nearest
neighbor list.

A further modification suggested by Schnitzer and co-workers was to use the
average distance to multiple nearest neighbors to calculate $\sigma_i$, a
technique used by [@jegou2007contextual].

In the field of dimensionality reduction, generating a nearest neighbor graph
is a step for many methods. [@JMLR:v22:20-1061] describe using local scaling
as a post-processing step with the following recipe:

* Find the $k+50$th nearest neighbors, where $k$ is the actual number of
neighbors you want (e.g. if you want to 15 neighbors per point, find the 65
nearest neighbors).
* Use the mean distance to the 4-6th nearest neighbors as $\sigma_i$. Using
this range was also advocated in another dimensionality reduction method [@DBLP:journals/corr/abs-1910-00204].
* From the $k+50$ nearest neighbors, find the $k$ neighbors with the smallest
locally scaled distances. These are the $k$ neighbors to use for downstream
processing.

To do with our 1000D example, we first need to generate the (approximate)
65-nearest neighbors:

```{r 65 nearest neighbors}
g1000d_nnd65 <- nnd_knn(g1000d, k = 65, metric = "euclidean")
```

This is a noticeably slow step. Next, we use the `local_scale_nn` function
to extract a 15-neighbor locally-scaled subset:

```{r local scale}
g1000d_nnd65ls15 <-
  local_scale_nn(g1000d_nnd65, k = 15, k_scale = c(4, 6))
```

If you look at the distances in `g1000d_nnd65ls15` these are the original
unscaled distances, as found in `g1000d_nnd65` ($d_{ij}$). The scaled distances
$\hat{d}_{ij}$ are used internally to select the 15 neighbors, but are not
returned from `local_scale_nn`.

How does local scaling affect the hubness of the resulting neighbor graph?

```{r hubness of locally scaled nearest neighbor graph}
g1000d_lsko <- k_occur(g1000d_nnd65ls15)
hist(g1000d_lsko, main = "locally scaled k-occ", xlab = "k-occurrences")
summary(g1000d_lsko)
sum(g1000d_lsko == 1)
```

This is a massive reduction in hubness over the 15NN graph, and nearly as good
as the search graph generated from the 15NN graph. Can we see further
improvements if we generate a search graph from the locally scaled neighbors?

```{r prepare search graph from locally scaled neighbors}
g1000dls_search_graph <-
  prepare_search_graph(
    data = g1000d,
    graph = g1000d_nnd65ls15,
    metric = "euclidean",
    diversify_prob = 1,
    pruning_degree_multiplier = 1.5
  )
```

```{r hubness of locally scaled search graph}
g1000d_lssgko <- k_occur(g1000dls_search_graph)
hist(g1000d_lssgko, main = "locally scaled search k-occ", xlab = "k-occurrences")
summary(g1000d_lssgko)
sum(g1000d_lssgko == 1)
```

Here the hubness properties are further improved over using either local scaling
or creating a search graph separately. While this is promising, does it
translate into better search performance? Here are the accuracy results using
the locally scaled search graph without any backtracking:

```{r search with locally scaled search graph, no back-tracking}
g1000dls_search0nnd <-
    graph_knn_query(
        query = g1000d,
        reference = g1000d,
        reference_graph = g1000dls_search_graph,
        k = 15,
        metric = "euclidean",
        init = g1000d_nnd,
        epsilon = 0
    )
nn_accuracy(g1000d_nnbf, g1000dls_search0nnd, k = 15)
```

Without having to do any back-tracking, we can get 94% accuracy, a nearly 10%
improvement compared to the search graph generated from the 15NN graph 
directly.

It's worth noting here that we initialized the search using `g1000d_nnd`, the
15NN results from running NND with `k = 15`. This makes comparing the
performance with the other search graph fairer, but in reality we wouldn't
have that data to hand, as we used `k = 65`. As noted in the 'Use More Neighbors'
section above, we already know that using NND with an increased value of `k` is
going to result in a big improvement without bothering with local scaling:

```{r NND 1000D truncated 65}
nn_accuracy(g1000d_nnbf, g1000d_nnd65, k = 15)
```

Almost perfect. We also have the locally-scaled 15 nearest neighbor graph we
can use for initialization:

```{r search with locally scaled search graph, no back-tracking, LS-initialized}
g1000dls_search0 <-
    graph_knn_query(
        query = g1000d,
        reference = g1000d,
        reference_graph = g1000dls_search_graph,
        k = 15,
        metric = "euclidean",
        init = g1000d_nnd65ls15,
        epsilon = 0
    )
nn_accuracy(g1000d_nnbf, g1000dls_search0, k = 15)
```

This is noticeably less good than initializing using the 15 NND results, but
it's still the best non-back-tracking result. If we do set `epsilon = 0.1`, then
we get 100% accuracy using both the neighbor graph and the search graph, but
then again the other non-locally scaled graphs are getting very close to that
too.

The value of local scaling is probably not if you want to find an accurate
approximation to the nearest neighbors: using back-tracking search with a search
graph derived from the unscaled neighbor graph directly is going to be less
effort. However there may be cases where the increased cost of building the
$k+50$ nearest neighbor graph is not a big problem, e.g. if you intend to build
a graph once but search it many times, then the search graph built from the
locally scaled neighbor graph is probably more efficient to traverse in cases
where the unscaled data suffers from hubs. Or the properties of the
locally-scaled graph may be more attractive for other uses e.g. dimensionality
reduction methods like those described by Wang and co-workers.

## Conclusions

* High dimensional data leads to hubs.
* The "hubness" of an item in a dataset can be measured by the k-occurrence in
the corresponding nearest neighbor graph. The higher the k-occurrence, the more
of a hub it is.
* The existence of a hubs implies the existence of "anti-hubs", i.e. items with
a low k-occurrence. A small number of hubs can create a disproportionately
larger number of anti-hubs, with a larger value of the k-occurrence creating 
more anti-hubs.
* The accuracy of nearest neighbor descent is reduced by the presence of hubs: 
specifically, the lower the k-occurrence of an item, the greater the probability
of a low accuracy of its nearest neighbors.
* Accuracy of nearest neighbor descent can be improved by searching for a larger
number of neighbors and then truncating the result to the desired size, at the
cost of a longer run-time and memory usage.
* Alternatively, you can run the the nearest neighbor descent multiple times
from different random starting points and merge the results.
* More accurate and efficient results are obtained by converting the nearest
neighbor descent results into a search graph and then querying the graph with
the original data, using the nearest neighbor results for initialization and
a back-tracking search.
* Local scaling can also reduce the effect of hubs and be combined with a search
graph, at the cost of a larger computational effort.

If you are concerned with potential hubs interfering with the accuracy of the
neighbor graph, I suggest the following steps:

1. Generate a neighbor graph with `nnd_knn` and default parameters.
2. Evaluate the hubness of the graph with `k_occur`.
3. If the maximum k-occurrence exceeds a threshold (maybe `10 * k` is a good
starting point), then use `prepare_search_graph` and `graph_knn_query` with 
back-tracking search (set `epsilon > 0`) to refine the results further.

This should provide a robust approach to producing accurate approximate
nearest neighbors without spending time on unnecessary graph search when the
results are probably already quite good.

For more on the effect of hubness and nearest neighbors, and more advanced attempts
to fix the problem, see the work of Flexer and co-workers
[@schnitzer2012local; @flexer2016empirical; @feldbauer2019comprehensive; @feldbauer2019scikit] 
and Radovanović and co-workers [@radovanovic2010hubs; @bratic2019influence].

## References
